---
title: "PTM Enrichment"
format:
  html:
    fig-dpi: 96
    toc-location: left
    grid:
      sidebar-width: 300px
    include-after-body: fgcz-plot-finder.html
knitr:
  opts_chunk:
    fig.retina: 1
params:
  json_path: null
  analysis_label: null
  fdr_threshold: 0.1
  top_terms: 10
  running_terms: 1
vignette: >
  %\VignetteIndexEntry{PTM Enrichment}
  %\VignetteEngine{quarto::format}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
#| include: false
required_packages <- c(
  "clusterProfiler",
  "dplyr",
  "DT",
  "enrichplot",
  "fgczQuartoTemplate",
  "ggplot2",
  "ggridges",
  "jsonlite",
  "knitr",
  "patchwork",
  "protsea"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, quietly = TRUE, FUN.VALUE = logical(1))
]
if (length(missing_packages)) {
  stop(
    "Missing package(s) required to render this report: ",
    paste(missing_packages, collapse = ", "),
    call. = FALSE
  )
}

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5.2,
  out.width = "92%",
  fig.align = "center"
)

input_candidates <- c(
  params$json_path,
  system.file("extdata", "PTM_DPA_gsea_result.json.gz", package = "protsea"),
  file.path("..", "inst", "extdata", "PTM_DPA_gsea_result.json.gz"),
  file.path("inst", "extdata", "PTM_DPA_gsea_result.json.gz")
)
input_candidates <- input_candidates[
  !is.na(input_candidates) & nzchar(input_candidates) & file.exists(input_candidates)
]
if (!length(input_candidates)) {
  stop("PTM GSEA JSON input was not found.", call. = FALSE)
}
json_path <- normalizePath(input_candidates[[1L]], mustWork = TRUE)
payload <- jsonlite::fromJSON(json_path, simplifyVector = FALSE)

abstract_candidates <- c(
  system.file("report-assets", "ptm-enrichment.svg", package = "protsea"),
  file.path("..", "inst", "report-assets", "ptm-enrichment.svg"),
  file.path("inst", "report-assets", "ptm-enrichment.svg")
)
abstract_candidates <- abstract_candidates[
  nzchar(abstract_candidates) & file.exists(abstract_candidates)
]
if (!length(abstract_candidates)) {
  stop("PTM enrichment visual abstract was not found.", call. = FALSE)
}
abstract_path <- normalizePath(abstract_candidates[[1L]], mustWork = TRUE)

required_top_level <- c("data", "rank_lists")
missing_top_level <- setdiff(required_top_level, names(payload))
if (length(missing_top_level)) {
  stop(
    "PTM GSEA JSON is missing: ",
    paste(missing_top_level, collapse = ", "),
    call. = FALSE
  )
}

category_names <- c(
  PTMSEA = "PTM-SEA",
  KinaseGSEA = "KinaseLib",
  MEA = "MEA"
)
method_labels <- c(
  PTMSEA = "PTM-SEA",
  KinaseGSEA = "Kinase GSEA",
  MEA = "MEA"
)
expected_categories <- unname(category_names)
document_categories <- unique(unlist(lapply(
  payload$data,
  function(contrast) names(contrast$categories)
)))
missing_categories <- setdiff(expected_categories, document_categories)
unexpected_categories <- setdiff(document_categories, expected_categories)
if (length(missing_categories) || length(unexpected_categories)) {
  stop(
    "Expected exactly the PTM-SEA, KinaseLib, and MEA categories; missing: ",
    paste(missing_categories, collapse = ", "),
    "; unexpected: ",
    paste(unexpected_categories, collapse = ", "),
    call. = FALSE
  )
}

decoded <- protsea::decode_gsea_json(json_path)
analysis_label <- params$analysis_label
if (is.null(analysis_label) || !length(analysis_label) || !nzchar(analysis_label)) {
  analysis_label <- payload$metadata$analysis
}
if (is.null(analysis_label) || !length(analysis_label) || !nzchar(analysis_label)) {
  analysis_label <- "PTM analysis"
}
analysis_label <- as.character(analysis_label[[1L]])

fdr_threshold <- as.numeric(params$fdr_threshold)
top_terms <- as.integer(params$top_terms)
running_terms <- as.integer(params$running_terms)
if (length(fdr_threshold) != 1L || !is.finite(fdr_threshold) ||
    fdr_threshold <= 0 || fdr_threshold > 1) {
  stop("fdr_threshold must be one number in (0, 1].", call. = FALSE)
}
if (length(top_terms) != 1L || is.na(top_terms) || top_terms < 1L) {
  stop("top_terms must be one positive integer.", call. = FALSE)
}
if (length(running_terms) != 1L || is.na(running_terms) || running_terms < 1L) {
  stop("running_terms must be one positive integer.", call. = FALSE)
}

empty_terms <- function() {
  data.frame(
    contrast = character(),
    term = character(),
    NES = numeric(),
    FDR = numeric(),
    direction = character(),
    mapped_sites = integer(),
    set_size = integer(),
    leading_edge_sites = character(),
    stringsAsFactors = FALSE
  )
}

scalar_number <- function(value) {
  if (is.null(value) || !length(value)) {
    return(NA_real_)
  }
  as.numeric(unlist(value, use.names = FALSE)[[1L]])
}

term_table_from_payload <- function(category) {
  tables <- lapply(names(payload$data), function(contrast_name) {
    terms <- payload$data[[contrast_name]]$categories[[category]]$terms
    if (!length(terms)) {
      return(empty_terms())
    }
    do.call(rbind, lapply(terms, function(term) {
      data.frame(
        contrast = contrast_name,
        term = as.character(term$description),
        NES = scalar_number(term$enrichment_score),
        FDR = scalar_number(term$fdr),
        direction = as.character(term$direction),
        mapped_sites = as.integer(scalar_number(term$genes_mapped)),
        set_size = as.integer(scalar_number(term$genes_in_set)),
        leading_edge_sites = paste(
          as.character(unlist(term$leading_edge_ids, use.names = FALSE)),
          collapse = ";"
        ),
        stringsAsFactors = FALSE
      )
    }))
  })
  dplyr::bind_rows(tables)
}

read_method_data <- function(method) {
  category <- category_names[[method]]
  native <- lapply(decoded, function(contrast) contrast[[category]])
  list(
    method = method,
    label = method_labels[[method]],
    category = category,
    contrasts = names(payload$data),
    table = term_table_from_payload(category),
    native = native
  )
}

enrichment <- stats::setNames(
  lapply(names(method_labels), read_method_data),
  names(method_labels)
)

summary_table <- function(data) {
  dplyr::bind_rows(lapply(data$contrasts, function(contrast_name) {
    current <- data$table[data$table$contrast == contrast_name, , drop = FALSE]
    data.frame(
      Contrast = gsub("_", " ", contrast_name, fixed = TRUE),
      `Tested terms` = nrow(current),
      `FDR-significant` = sum(current$FDR <= fdr_threshold, na.rm = TRUE),
      `Positive NES` = sum(
        current$FDR <= fdr_threshold & current$NES > 0,
        na.rm = TRUE
      ),
      `Negative NES` = sum(
        current$FDR <= fdr_threshold & current$NES < 0,
        na.rm = TRUE
      ),
      check.names = FALSE
    )
  }))
}

summary_widget <- function(data) {
  DT::datatable(
    summary_table(data),
    filter = "none",
    rownames = FALSE,
    class = "compact stripe",
    options = list(pageLength = 10, scrollX = TRUE, dom = "t")
  )
}

empty_plot <- function(label) {
  ggplot2::ggplot() +
    ggplot2::annotate("text", x = 0, y = 0, label = label) +
    ggplot2::theme_void()
}

nonempty_native_results <- function(data) {
  data$native[vapply(
    data$native,
    function(result) nrow(result@result) > 0L,
    logical(1)
  )]
}

combined_native_result <- function(data) {
  usable <- nonempty_native_results(data)
  if (!length(usable)) {
    return(NULL)
  }
  if (length(usable) == 1L) {
    return(usable[[1L]])
  }
  clusterProfiler::merge_result(usable)
}

dot_plot <- function(data) {
  object <- combined_native_result(data)
  if (is.null(object)) {
    return(empty_plot("No enrichment terms are available."))
  }
  enrichplot::dotplot(object, showCategory = top_terms) +
    ggplot2::theme(
      axis.text.y = ggplot2::element_text(size = 8),
      legend.position = "bottom"
    )
}

heatmap_plot <- function(data) {
  if (!nrow(data$table)) {
    return(empty_plot("No enrichment terms are available."))
  }
  filter_threshold <- if (any(data$table$FDR <= 0.15, na.rm = TRUE)) 0.15 else 1
  top_items <- data$table |>
    dplyr::filter(.data$FDR < filter_threshold) |>
    dplyr::group_by(.data$term) |>
    dplyr::summarize(min_fdr = min(.data$FDR), .groups = "drop") |>
    dplyr::arrange(.data$min_fdr) |>
    utils::head(top_terms) |>
    dplyr::pull(.data$term)
  plot_data <- data$table |>
    dplyr::filter(.data$term %in% top_items) |>
    dplyr::mutate(
      significance = dplyr::case_when(
        .data$FDR < 0.01 ~ "***",
        .data$FDR < 0.05 ~ "**",
        .data$FDR < 0.1 ~ "*",
        TRUE ~ ""
      )
    )
  ggplot2::ggplot(
    plot_data,
    ggplot2::aes(
      x = .data$contrast,
      y = stats::reorder(.data$term, .data$NES),
      fill = .data$NES
    )
  ) +
    ggplot2::geom_tile(colour = "white") +
    ggplot2::geom_text(ggplot2::aes(label = .data$significance), size = 4) +
    ggplot2::scale_fill_gradient2(low = "#2b6cb0", mid = "white", high = "#c53030") +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
      axis.text.y = ggplot2::element_text(size = 8)
    ) +
    ggplot2::labs(
      x = NULL,
      y = NULL,
      fill = "NES",
      subtitle = "* FDR < 0.1, ** FDR < 0.05, *** FDR < 0.01"
    )
}

volcano_plot <- function(data) {
  volcano_data <- data$table |>
    dplyr::filter(is.finite(.data$NES), is.finite(.data$FDR)) |>
    dplyr::mutate(
      negative_log10_fdr = -log10(pmax(.data$FDR, 1e-10)),
      direction = dplyr::case_when(
        .data$NES > 0 ~ "Up",
        .data$NES < 0 ~ "Down",
        TRUE ~ "NS"
      ),
      significant = .data$FDR < fdr_threshold
    )
  if (!nrow(volcano_data)) {
    return(empty_plot("No finite enrichment statistics are available."))
  }
  label_data <- volcano_data |>
    dplyr::filter(.data$FDR < fdr_threshold) |>
    dplyr::group_by(.data$contrast) |>
    dplyr::slice_min(.data$FDR, n = min(5L, top_terms)) |>
    dplyr::ungroup() |>
    dplyr::mutate(label_hjust = ifelse(.data$NES < 0, 1.1, -0.1))
  ggplot2::ggplot(
    volcano_data,
    ggplot2::aes(x = .data$NES, y = .data$negative_log10_fdr)
  ) +
    ggplot2::geom_point(
      ggplot2::aes(colour = .data$direction, alpha = .data$significant),
      size = 2
    ) +
    ggplot2::geom_hline(
      yintercept = -log10(fdr_threshold),
      linetype = "dashed",
      colour = "grey30"
    ) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", colour = "grey30") +
    ggplot2::geom_text(
      data = label_data,
      ggplot2::aes(label = .data$term, hjust = .data$label_hjust),
      size = 2.5,
      check_overlap = TRUE
    ) +
    ggplot2::scale_x_continuous(
      expand = ggplot2::expansion(mult = c(0.16, 0.16))
    ) +
    ggplot2::scale_colour_manual(
      values = c(Up = "red", Down = "blue", NS = "grey50")
    ) +
    ggplot2::scale_alpha_manual(
      values = c(`TRUE` = 1, `FALSE` = 0.3),
      guide = "none"
    ) +
    ggplot2::facet_wrap(~ .data$contrast, scales = "free") +
    ggplot2::theme_bw() +
    ggplot2::theme(
      legend.position = "bottom",
      plot.margin = ggplot2::margin(5.5, 18, 5.5, 18)
    ) +
    ggplot2::labs(
      subtitle = paste0(
        "Dashed line: FDR = ",
        fdr_threshold,
        "; labels: top ",
        min(5L, top_terms),
        " by FDR per contrast"
      ),
      x = "NES",
      y = "-log10(FDR)",
      colour = "Direction"
    )
}

running_score_plots <- function(data) {
  plots <- list()
  for (contrast_name in names(data$native)) {
    result <- data$native[[contrast_name]]
    if (!nrow(result@result)) {
      next
    }
    top <- result@result |>
      dplyr::arrange(.data$p.adjust, .data$pvalue) |>
      utils::head(running_terms)
    for (i in seq_len(nrow(top))) {
      term_id <- top$ID[[i]]
      title <- paste0(
        gsub("_", " ", contrast_name, fixed = TRUE),
        " · ",
        top$Description[[i]],
        " (NES=", round(top$NES[[i]], 2),
        ", FDR=", signif(top$p.adjust[[i]], 2), ")"
      )
      plots[[paste(contrast_name, term_id, sep = "__")]] <- enrichplot::gseaplot2(
        result,
        geneSetID = term_id,
        title = title,
        pvalue_table = FALSE
      )
    }
  }
  plots
}

as_patchwork <- function(plot) {
  if (!inherits(plot, "gglist")) {
    return(plot)
  }
  params <- attr(plot, "params")
  patchwork::wrap_plots(
    unclass(plot),
    ncol = params$ncol,
    heights = params$heights
  )
}

print_running_scores <- function(data) {
  plots <- running_score_plots(data)
  if (!length(plots)) {
    cat("::: {.callout-note}\nNo enrichment term is available for a running-score plot.\n:::\n")
    return(invisible(NULL))
  }
  plots <- lapply(plots, as_patchwork)
  print(patchwork::wrap_plots(plots, ncol = min(2L, length(plots))))
  invisible(NULL)
}

rank_distribution_plots <- function(data) {
  usable <- nonempty_native_results(data)
  lapply(names(usable), function(contrast_name) {
    result <- usable[[contrast_name]]
    enrichplot::ridgeplot(
      result,
      showCategory = as.numeric(min(top_terms, nrow(result@result))),
      fill = "p.adjust"
    ) +
      ggplot2::labs(
        title = paste0(
          gsub("_", " ", contrast_name, fixed = TRUE),
          " · ranked-site distributions"
        ),
        x = "Site ranking statistic",
        fill = "FDR"
      ) +
      ggplot2::theme(legend.position = "bottom")
  })
}

print_rank_distributions <- function(data) {
  plots <- rank_distribution_plots(data)
  if (!length(plots)) {
    cat("::: {.callout-note}\nNo enrichment term is available for a rank-distribution plot.\n:::\n")
    return(invisible(NULL))
  }
  print(patchwork::wrap_plots(plots, ncol = min(2L, length(plots))))
  invisible(NULL)
}

gene_set_network_plot <- function(data) {
  object <- combined_native_result(data)
  if (is.null(object)) {
    return(empty_plot("No enrichment term is available for a gene-set network."))
  }
  enrichplot::cnetplot(
    object,
    showCategory = min(4L, top_terms),
    node_label = "category"
  ) +
    ggplot2::labs(
      title = paste0(analysis_label, " · ", data$label),
      subtitle = "Leading terms and their shared ranked sites"
    )
}

term_similarity_plots <- function(data) {
  object <- combined_native_result(data)
  if (is.null(object)) {
    return(list())
  }
  result <- if (methods::is(object, "compareClusterResult")) {
    object@compareClusterResult
  } else {
    object@result
  }
  if (length(unique(result$ID)) < 3L) {
    return(list())
  }
  n_show <- min(12L, top_terms, length(unique(result$ID)))
  similarity <- enrichplot::pairwise_termsim(object)
  list(
    enrichplot::emapplot(similarity, showCategory = n_show) +
      ggplot2::labs(
        title = "Enrichment map",
        subtitle = "Terms are connected by ranked-site overlap"
      ),
    enrichplot::treeplot(similarity, showCategory = n_show) +
      ggplot2::labs(
        title = "Term-similarity tree",
        subtitle = "Terms are grouped by ranked-site overlap"
      )
  )
}

print_term_similarity <- function(data) {
  plots <- term_similarity_plots(data)
  if (!length(plots)) {
    cat("::: {.callout-note}\nAt least three enrichment terms are required for term-similarity plots.\n:::\n")
    return(invisible(NULL))
  }
  print(patchwork::wrap_plots(plots, ncol = min(2L, length(plots))))
  invisible(NULL)
}

result_widget <- function(data) {
  display <- data$table |>
    dplyr::transmute(
      Contrast = gsub("_", " ", .data$contrast, fixed = TRUE),
      Term = .data$term,
      NES = round(.data$NES, 3),
      FDR = signif(.data$FDR, 3),
      Direction = .data$direction,
      `Mapped sites` = .data$mapped_sites,
      `Set size` = .data$set_size,
      `Leading-edge sites` = .data$leading_edge_sites
    ) |>
    dplyr::arrange(.data$Contrast, .data$FDR)
  DT::datatable(
    display,
    filter = "top",
    rownames = FALSE,
    class = "compact stripe",
    extensions = "Buttons",
    options = list(
      pageLength = 12,
      scrollX = TRUE,
      scrollY = "420px",
      deferRender = TRUE,
      dom = "Bfrtip",
      buttons = c("copy", "csv", "excel")
    )
  )
}

overview <- dplyr::bind_rows(lapply(enrichment, function(method) {
  data.frame(
    Tool = method$label,
    Contrasts = length(method$contrasts),
    Terms = nrow(method$table),
    `FDR-significant` = sum(method$table$FDR <= fdr_threshold, na.rm = TRUE),
    check.names = FALSE
  )
}))
generated_at <- format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")
```

```{r report-metadata}
#| results: asis
cat(sprintf(
  '<div id="fgcz-report-metadata" hidden data-generated-at="%s"></div>',
  generated_at
))
```

::: {.panel-tabset}

# Overview

<img src="`r knitr::image_uri(abstract_path)`" alt="Visual abstract showing DPA, DPU, and CorrectFirst rankings entering PTM-SEA, Kinase GSEA, and MEA, followed by comparable enrichment plots and result tables decoded from JSON." style="width: 100%; max-height: 20rem; object-fit: contain;">

This report visualizes one completed **`r analysis_label`** enrichment document. The document contains a shared ranked list for every contrast and the results from **PTM-SEA**, **Kinase GSEA**, and **MEA**. Every statistic, membership, leading edge, and running enrichment score is decoded from the supplied JSON; this report does not rerun enrichment.

```{r}
#| label: tbl-tool-overview
#| tbl-cap: "PTM enrichment tools, contrasts, tested terms, and terms passing the report FDR threshold."
DT::datatable(
  overview,
  filter = "none",
  rownames = FALSE,
  class = "compact stripe",
  options = list(pageLength = 10, scrollX = TRUE, dom = "t")
)
```

# PTM-SEA

PTM-SEA tests curated kinase, perturbation, pathway, and disease-associated PTM signatures against the `r analysis_label` site ranking.

::: {.panel-tabset}

## Summary

```{r}
#| label: tbl-ptmsea-summary
#| tbl-cap: "PTM-SEA result counts by contrast and enrichment direction."
summary_widget(enrichment$PTMSEA)
```

## Dot plot

```{r}
#| label: fig-ptmsea-dotplot
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "94%"
#| fig-cap: "PTM-SEA normalized enrichment scores by contrast. Point position gives NES, colour gives adjusted significance, and point size gives the mapped-site count."
dot_plot(enrichment$PTMSEA)
```

## Heatmap

```{r}
#| label: fig-ptmsea-heatmap
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "94%"
#| fig-cap: "PTM-SEA normalized enrichment scores for the leading signatures across contrasts. Tile colour gives NES and symbols indicate FDR thresholds."
heatmap_plot(enrichment$PTMSEA)
```

## Volcano

```{r}
#| label: fig-ptmsea-volcano
#| fig-width: 12
#| fig-height: 4.4
#| out-width: "96%"
#| fig-cap: "PTM-SEA normalized enrichment score as a function of FDR. Each point is one signature and each panel is one contrast."
volcano_plot(enrichment$PTMSEA)
```

## Running score

```{r}
#| label: fig-ptmsea-running
#| fig-width: 12
#| fig-height: 5.8
#| out-width: "96%"
#| fig-cap: "Running enrichment scores for the leading PTM-SEA signature in each contrast, reconstructed from the ordered ranks, signature membership, and weighting parameters stored in JSON."
#| results: asis
print_running_scores(enrichment$PTMSEA)
```

## Rank distributions

```{r}
#| label: fig-ptmsea-ridge
#| fig-width: 12
#| fig-height: 5.4
#| out-width: "96%"
#| fig-cap: "Ranked-site distributions for the leading PTM-SEA signatures. Each ridge shows the ranking statistics of sites belonging to one signature and colour encodes FDR."
#| results: asis
print_rank_distributions(enrichment$PTMSEA)
```

## Gene-set network

```{r}
#| label: fig-ptmsea-network
#| fig-width: 11
#| fig-height: 6.2
#| out-width: "94%"
#| fig-cap: "PTM-SEA signature network. Signature nodes connect to their member sites, exposing shared sites across leading terms and contrasts."
gene_set_network_plot(enrichment$PTMSEA)
```

## Term similarity

```{r}
#| label: fig-ptmsea-similarity
#| fig-width: 12
#| fig-height: 5.8
#| out-width: "96%"
#| fig-cap: "PTM-SEA enrichment map and similarity tree. Signatures are related by overlap among their ranked-site members."
#| results: asis
print_term_similarity(enrichment$PTMSEA)
```

## Results

```{r}
#| label: tbl-ptmsea-results
#| tbl-cap: "Complete PTM-SEA result table decoded from the shared JSON document."
result_widget(enrichment$PTMSEA)
```

:::

# Kinase GSEA

Kinase GSEA tests predicted kinase-substrate sets against the same `r analysis_label` ranked sites.

::: {.panel-tabset}

## Summary

```{r}
#| label: tbl-kinase-summary
#| tbl-cap: "Kinase GSEA result counts by contrast and enrichment direction."
summary_widget(enrichment$KinaseGSEA)
```

## Dot plot

```{r}
#| label: fig-kinase-dotplot
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "94%"
#| fig-cap: "Kinase GSEA normalized enrichment scores by contrast. Point position gives NES, colour gives adjusted significance, and point size gives the mapped-site count."
dot_plot(enrichment$KinaseGSEA)
```

## Heatmap

```{r}
#| label: fig-kinase-heatmap
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "94%"
#| fig-cap: "Kinase GSEA normalized enrichment scores for the leading kinases across contrasts. Tile colour gives NES and symbols indicate FDR thresholds."
heatmap_plot(enrichment$KinaseGSEA)
```

## Volcano

```{r}
#| label: fig-kinase-volcano
#| fig-width: 12
#| fig-height: 4.4
#| out-width: "96%"
#| fig-cap: "Kinase GSEA normalized enrichment score as a function of FDR. Each point is one kinase and each panel is one contrast."
volcano_plot(enrichment$KinaseGSEA)
```

## Running score

```{r}
#| label: fig-kinase-running
#| fig-width: 12
#| fig-height: 5.8
#| out-width: "96%"
#| fig-cap: "Running enrichment scores for the leading kinase set in each contrast, reconstructed from the ordered ranks, substrate membership, and weighting parameters stored in JSON."
#| results: asis
print_running_scores(enrichment$KinaseGSEA)
```

## Rank distributions

```{r}
#| label: fig-kinase-ridge
#| fig-width: 12
#| fig-height: 5.4
#| out-width: "96%"
#| fig-cap: "Ranked-site distributions for the leading kinase substrate sets. Each ridge shows the ranking statistics of sites assigned to one kinase and colour encodes FDR."
#| results: asis
print_rank_distributions(enrichment$KinaseGSEA)
```

## Gene-set network

```{r}
#| label: fig-kinase-network
#| fig-width: 11
#| fig-height: 6.2
#| out-width: "94%"
#| fig-cap: "Kinase substrate-set network. Kinase nodes connect to their predicted substrate sites, exposing shared sites across leading kinases and contrasts."
gene_set_network_plot(enrichment$KinaseGSEA)
```

## Term similarity

```{r}
#| label: fig-kinase-similarity
#| fig-width: 12
#| fig-height: 5.8
#| out-width: "96%"
#| fig-cap: "Kinase GSEA enrichment map and similarity tree. Kinases are related by overlap among their predicted substrate sites."
#| results: asis
print_term_similarity(enrichment$KinaseGSEA)
```

## Results

```{r}
#| label: tbl-kinase-results
#| tbl-cap: "Complete Kinase GSEA result table decoded from the shared JSON document."
result_widget(enrichment$KinaseGSEA)
```

:::

# MEA

MEA reports the native GSEApy kinase-motif enrichment against the same `r analysis_label` ranked sites. The shared JSON retains its source running scores and hit positions.

::: {.panel-tabset}

## Summary

```{r}
#| label: tbl-mea-summary
#| tbl-cap: "MEA result counts by contrast and enrichment direction."
summary_widget(enrichment$MEA)
```

## Dot plot

```{r}
#| label: fig-mea-dotplot
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "94%"
#| fig-cap: "MEA normalized enrichment scores by contrast. Point position gives native NES, colour gives adjusted significance, and point size gives the mapped-site count."
dot_plot(enrichment$MEA)
```

## Heatmap

```{r}
#| label: fig-mea-heatmap
#| fig-width: 11
#| fig-height: 5.2
#| out-width: "94%"
#| fig-cap: "MEA normalized enrichment scores for the leading kinase motifs across contrasts. Tile colour gives NES and symbols indicate FDR thresholds."
heatmap_plot(enrichment$MEA)
```

## Volcano

```{r}
#| label: fig-mea-volcano
#| fig-width: 12
#| fig-height: 4.4
#| out-width: "96%"
#| fig-cap: "MEA normalized enrichment score as a function of FDR. Each point is one kinase motif and each panel is one contrast."
volcano_plot(enrichment$MEA)
```

## Running score

```{r}
#| label: fig-mea-running
#| fig-width: 12
#| fig-height: 5.8
#| out-width: "96%"
#| fig-cap: "Native MEA running enrichment scores for the leading kinase motif in each contrast, reconstructed from the GSEApy scores and hit positions stored in JSON."
#| results: asis
print_running_scores(enrichment$MEA)
```

## Rank distributions

```{r}
#| label: fig-mea-ridge
#| fig-width: 12
#| fig-height: 5.4
#| out-width: "96%"
#| fig-cap: "Ranked-site distributions for the leading MEA kinase motifs. Each ridge shows the ranking statistics of substrate sites and colour encodes FDR."
#| results: asis
print_rank_distributions(enrichment$MEA)
```

## Gene-set network

```{r}
#| label: fig-mea-network
#| fig-width: 11
#| fig-height: 6.2
#| out-width: "94%"
#| fig-cap: "MEA kinase-motif network. Kinase nodes connect to their substrate sites, exposing shared sites across leading motifs and contrasts."
gene_set_network_plot(enrichment$MEA)
```

## Term similarity

```{r}
#| label: fig-mea-similarity
#| fig-width: 12
#| fig-height: 5.8
#| out-width: "96%"
#| fig-cap: "MEA enrichment map and similarity tree. Kinase motifs are related by overlap among their substrate sites."
#| results: asis
print_term_similarity(enrichment$MEA)
```

## Results

```{r}
#| label: tbl-mea-results
#| tbl-cap: "Complete MEA result table decoded from the shared JSON document."
result_widget(enrichment$MEA)
```

:::

# Session Info

::: {.panel-tabset}

## Report provenance

```{r}
#| label: tbl-report-provenance
#| tbl-cap: "Input document, report parameters, package version, and render timestamp."
qmd_source <- fgczQuartoTemplate::fgcz_qmd_source_link()
provenance <- data.frame(
  Field = c(
    "Analysis",
    "Input JSON",
    "Contrasts",
    "Tools",
    "FDR threshold",
    "Top terms per summary plot",
    "Running-score terms per contrast",
    "Report package",
    "Generated at",
    "Qmd source"
  ),
  Value = c(
    analysis_label,
    json_path,
    paste(names(payload$data), collapse = "; "),
    paste(unname(method_labels), collapse = "; "),
    format(fdr_threshold),
    top_terms,
    running_terms,
    paste0("protsea ", utils::packageVersion("protsea")),
    generated_at,
    qmd_source
  ),
  stringsAsFactors = FALSE
)
DT::datatable(
  provenance,
  filter = "none",
  rownames = FALSE,
  escape = FALSE,
  class = "compact stripe",
  options = list(pageLength = 10, scrollX = TRUE, dom = "t")
)
```

## R session info

```{r session-info}
utils::sessionInfo()
```

:::

:::
